AGB Stars in the IR: Physical Characteristics, Galactic Spatial Distribution, and Chemical Impact of the Milky Way AGB Population

### 2015-03-31 paper1/paper1_criteria.tex
Updating paper

\section{Object Selection Criteria}
\label{sec:criteria}
We made color-color criteria to create a catalog of AGB candidates. We kept it simple to maximize the number of objects we'd retain while minimizing contamination from non-AGB objects.

>>>>

\section{Object Selection Criteria}
\label{sec:criteria}
We create color-color criteria to generate a catalog of AGB candidates. We seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.

The color-color criteria for AGB selection are as follows:
\begin{align} 
(J-K_s) > 1.1\label{eq:criteria1}\\
(W2-W3) > 0.3\label{eq:criteria2}\\
(W3-W4) < -0.83(W2-W3) + 3.37\label{eq:criteria3}
\end{align}
The criteria in (1) and (2) are primarily concerned with rejecting objects from the stellar locus, and other stars whose NIR spectra are dominated by the Rayleigh-Jeans tail (many covered by the sample of galaxies).  Criteria in (3) also reject stars from the stellar locus, but primarily function to reject IR-bright extragalactic sources.

\noindent Sample completeness $\eta$ is defined as
\begin{eqnarray*}
\eta &=& \frac{N - n_\text{missed}}{N}
\end{eqnarray*}
where $N$ is the total number of objects in that sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries. Figure~\ref{fig:completeness} shows the distribution of completeness amongst AGB sources after the application of the above criteria. Sample contamination is typically defined as
\begin{eqnarray*}
\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
\end{eqnarray*}
where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected}$ is the total number of selected sources, including AGB objects and spurious objects \citep{2013sdmm.book.....I}. The contamination map is shown in Figure~\ref{fig:contamination}.

\begin{figure}[h]
\includegraphics[width=6.7in]{figs/completeness_map.pdf}
\caption{Words \label{fig:completeness}.}
\end{figure}
\begin{figure}[h]
\includegraphics[width=6.7in]{figs/contamination_map.pdf}
\caption{Words \label{fig:contamination}.}
\end{figure}

A fairly large issue for creating the boundaries are the O-rich stars.  Either the objects from the OGLE O-rich AGB star catalog were mostly mis-matched to WISE, taking on colors of the stellar locus, or the issue is more physical.  It could be that the warm O-rich AGB photospheres are still visible and not heavily enshrouded in dust, thus appearing similar to Main Sequence stars in the NIR. Previous work on warm O-rich AGB stars has shown that their circumstellar shells are not prominent, and their NIR photometry reflects the Rayleigh Jeans tail of a cool 2000-4000K blackbody \citep{1974ApJ...189...89D}.

The results of the applied critiera are summarized in Table~\ref{tab:completeness}. Of the objects remaining in the sample, 52.94\% of the AGBs have $\lvert b\rvert \le 10^\circ$.

\begin{table}[h]
    \begin{center}
        \caption{Sample Completeness and Contamination}
        \scalebox{0.85}{
            \begin{tabular}{l c c c c c c}
		\hline
		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
		Completeness & 89.62\% & 72.11\% & 95.62\% & 39.53\% & 22.31\% \\ 
		\hline
		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
		Completeness & 88.45\% & 81.08\% & 28.77\% & 14.75\% \\ 
		\hline
		Population & OGLE-III C-rich & O-rich & \textbf{All AGB Stars} \\ 
		Completeness & 73.09\% & 70.68\% & \textbf{79.07\%} \\ 
		\hline
		Population & DR12 SSPP & DR7 LRG & Galaxies & QSO & AGN \\ 
		Contamination & 0.56\% & 0.00\% & 0.00\% & 0.07\% & 	0.00\% \\ 
		\hline

                \label{tab:completeness}
            \end{tabular}}    
    \end{center}
\end{table}

### 2015-03-31 paper1/paper1_data.tex
Updating paper

\subsection{Data Reduction}
We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise. We select only matches within 2" between our main sample of known \agb\, stars and \allwise. We further reduce our sample to unique matches to \allwise\, amongst \ogle, \macho, and \simbad\, by selecting \allwise\, photometry for unique \allwise\, designations. All samples of \agb\, star matches were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}. We require associations with \twomass\, objects within 3", detections in \twomass\, $J$, $K_s$, and all \allwise\, magnitudes as well as SNR $>$ 3 in each \allwise\, band.

We do not enforce the same faint or saturation limits on our contaminant samples. As a show of the strength of our selection criteria, we seek to include matches to each contaminant with reasonable detections in \allwise, and magnitudes reliable enough to perform the necessary color selections. We limit our sample from the \sdss\, DR9 SSPP to objects within 0.5 dex of the $g'-r'$ vs. $g'-i'$ stellar locus \citep{2014MNRAS.440.3430D}, \sdss\, SNR $>$ 5, [$W1/W2/W3/W4$] SNR $>$ 3,  detections in \twomass\, $J$ and $K_s$ bands, and one 2MASS association within 3". Objects from the NYU VAGC were reduced with requirements of SNR $>$ 3 for [$W1/W2/W3/W4$], required detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3". LRGs were required to have SNR $>$ 3 for [$W1/W2/W3$], detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3".  The restriction on $W4$ was dropped here, as it reduces the population of LRGs by an order of magnitude. The population for each sample from initial matching, as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}.\\

\vspace{-10pt}
\begin{table}[h]
	\begin{center}
	\caption{AGB and Contaminant Populations}
	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
		\hline
		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
		\hline
		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
		\hline\hline
		Population & MACHO seq1 & seq2 & seq3 & seq4\\
		\hline
		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
		Reduced & 927 & 642 & 263 & 336 \\
		\hline\hline
		Population & OGLE C-rich & O-rich\\
		\hline
		2" match & 11,417 & 38,369 \\
		Reduced & 737 & 2515 \\
		\hline\hline
		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
		\hline
		2" match & 1,508,158 & 18,481 & 102,178 & 113,392 & 799,761 \\
		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
		\hline

@@

>>>>

\subsection{Data Reduction}
We use NASA/IPAC IRSA's {\tt GATOR} tool\footnote{\url{http://irsa.ipac.caltech.edu/cgi-bin/Gator/nph-scan?mission=irsa&submit=Select&projshort=WISE}} to positionally match \sdss, \ogle, \macho, and \simbad\, to \allwise. We select only matches within 3" between our main sample of known \agb\, stars and \allwise. All samples of \agb\, star matches were required to be brighter than the published 5$\sigma$ faint limits of [16.83/15.6/11.32/8.0], as well as fainter than the saturation limits of [2.0/1.5/-3.0/-4.0] extrapolated from the wings of the PSFs for point sources, for [$W1/W2/W3/W4$] \citep{2013wise.rept....1C}, with no confusion or contamination flags in [$W1/W2/W3/W4$]. We require only single associations with \twomass\, objects within 3", detections in \twomass\, $J$, $K_s$, and all \allwise\, magnitudes as well as SNR $>$ 3 in each \allwise\, band. 
%We do not enforce the same faint or saturation limits on our contaminant samples. As a show of the strength of our selection criteria, we seek to include matches to each contaminant with reasonable detections in \allwise, and magnitudes reliable enough to perform the necessary color selections. We limit our sample from the \sdss\, DR9 SSPP to objects within 0.5 dex of the $g'-r'$ vs. $g'-i'$ stellar locus \citep{2014MNRAS.440.3430D}, \sdss\, SNR $>$ 5, [$W1/W2/W3/W4$] SNR $>$ 3,  detections in \twomass\, $J$ and $K_s$ bands, and one 2MASS association within 3". Objects from the NYU VAGC were reduced with requirements of SNR $>$ 3 for [$W1/W2/W3/W4$], required detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3". LRGs were required to have SNR $>$ 3 for [$W1/W2/W3$], detections for \twomass\, $J$ and $K_s$ bands, and one \twomass\, association within 3".  The restriction on $W4$ was dropped here, as it reduces the population of LRGs by an order of magnitude. 
The population for each sample from initial matching as well as after the application of the \allwise\, faint limits, saturation limits, and \twomass\, detection requirements are shown in Table~\ref{tab:pop}. The WISE color-color distributions for the AGB and contaminant samples are shown in Figure~\ref{fig:distros}. \\

\vspace{-10pt}
%\begin{table}[h]
%	\begin{center}
%	\caption{AGB and Contaminant Populations}
%	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
%		\hline
%		Population & SIMBAD C stars & OH/IR stars & Miras & S stars & AGB stars \\
%		\hline
%		2" match & 13,245 & 294 & 8,850 & 1,078 & 1,665 \\
%		Reduced & 3,327 & 165 & 6,218 & 865 & 1,121 \\
%		\hline\hline
%		Population & MACHO seq1 & seq2 & seq3 & seq4\\
%		\hline
%		2" match & 5,193 & 3,441 & 2,548 & 2,931 \\
%		Reduced & 927 & 642 & 263 & 336 \\
%		\hline\hline
%		Population & OGLE C-rich & O-rich\\
%		\hline
%		2" match & 11,417 & 38,369 \\
%		Reduced & 737 & 2515 \\
%		\hline\hline
%		Population & Locus Stars & AGN & LRG & QSO & Galaxies \\
%		\hline
%		2" match & 1,508,158 & 18,481 & 102,178 &  & 799,761 \\
%		Reduced & 168,045 & 9,652 & 7,717 & 18,360 & 125,869 \\
%		\hline
%
%		\label{tab:pop}
%	\end{tabular}}
%	\end{center}
%\end{table}

\begin{table}[h]
	\begin{center}
	\caption{AGB and Contaminant Populations}
	\scalebox{0.85}{\begin{tabular}{l c c c c c c}
		\hline\hline
		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
		\hline
		3" match & 1,689 & 14,209 & 9,027 & 406 & 1,081 \\
		Reduced & 684 & 1,782 & 3,241 & 43 & 511 \\ 
		\hline
		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
		\hline
		3" match & 5,279 & 3,519 & 2,619 & 3,070 \\
		Reduced & 277 & 185 & 73 & 61 \\ 
		\hline
		Population & OGLE-III C-rich & O-rich \\ 
		\hline
		3" match & 11,542 & 38,848 \\
		Reduced & 249 & 730 \\ 
		\hline
		Population & DR12 SSPP & DR7 LRG & QSO & AGN & Galaxies \\ 
		\hline
		3" match & 1,578,329 & 104,345 & 103,590 & 18,528 & 841,712 \\ 
		Reduced & 67,508 & 84 & 3,977 & 1,069 & 44,314 \\ 
		\hline\hline

@@

\begin{figure}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color1.pdf}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color2.pdf}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
\caption{\emph{Left:} Known AGB stars after data reduction. \emph{Right:} All contaminants after data reduction.\label{fig:distros}}
\end{figure}
